### 2015-05-13 papers/histogram/thesis/poster.tex
some changes to plots; added presentation

%%% math and physics
\usepackage{amsmath} % math packages
\renewcommand{\t}{\text} % for text in math environment
\usepackage[boldvectors,braket]{physymb} % physics packages
\renewcommand{\v}{\vec} % bold vectors
\newcommand{\p}[1]{\left(#1\right)} % bold vectors

%%% figures
\usepackage{graphicx,grffile,float,subcaption} % floats, etc.
\usepackage{multirow} % for multirow entries in tables

%% misc.
\usepackage[font=normalsize,labelfont=bf]{caption} % caption text

@@

\begin{poster}{
% options
    %% colors and appearance
    borderColor=border_color, %
    headerColorOne=header_color_left, %
    headerColorTwo=header_color_right, %
    headerFontColor=header_font_color, %
    boxColorOne=box_color, %
    textborder=rectangle, %
    headerborder=open, %
    background=none, %
    %% headers
    headerfont=\Large\sf\bf, %
    headerheight=0.11\textheight, %
    headershape=roundedright, %
    %% other
    columns=3 %
  }%
  {% top left logo
  }%
  {% title
    \LARGE\title
  }%
  {% authors
    \Large\author \\[1mm]
    \location
  }%
  {% top right logo
    % \includegraphics[width=2.5cm]{figs/osu.pdf}
    % \hspace{5mm}
  }%

@@

    \begin{itemize}
    \item Model system used to capture low order effects of
      short-range attractive forces, e.g. the van Der Waals force
    \item Composed of spheres with diameter $\sigma$ and the pair
      potential given in Figure \ref{fig:pair_potential}
    \item $\lambda$ is called the well width, and $\epsilon$ the well
      depth
    % \item Natural distance scale $\sigma$ and energy scale $\epsilon$
    \item First part of potential, $v_{sw}\p{r<\sigma}=\infty$,
      forbids spheres from overlapping
    \item Second part of potential,
      $v_{sw}\p{r:\sigma<r<\lambda\sigma}=-\epsilon$, associates an
      energy $-\epsilon$ with each pair of spheres whose centers are
      within distance of $\lambda\sigma$ of each other
    \item Spheres whose centers are more than $\lambda\sigma$ apart do
      not interact
    \end{itemize}

@@

    \begin{itemize}
    \item Model systems are powerful tools for understanding complex
      physical systems, but they do not exist in the real world
    \item Theories tested against Monte Carlo simulations
    \item Monte Carlo fluid simulations: moving spheres randomly to
      collect statistics on properties (e.g. potential energy, density,
      etc.) of valid fluid configurations
    \item Studying critical point requires sampling states with highly
      localized distribution of spheres, which is unlikely to occur by
      randomly sampling of all possible fluid configurations (second
      law of thermodynamics)
    \item For the square-well fluid, localized distribution of spheres
      is synonymous with a low energy state
    \end{itemize}

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.5\columnwidth]{figs/dos-example.pdf}
      \caption{Sample density of states of a particular square-well
        fluid ($\lambda=1.3$, $\eta=0.3$). The logarithmic span of the
        density of states is proportional to the number of moves one
        must simulate via standard Monte Carlo in order to observe the
        energies with the lowest state densities.}
      \label{fig:dos}
    \end{figure}

@@

    \begin{itemize}
    \item Means for biasing Monte Carlo simulations in favor of lower
      energy states
    \item Accept all moves in simulation which lower the system's
      energy, probabilistically reject moves to higher energy
    \item Details of rejection process define a histogram method
    \item There are several published histogram methods, but little to
      no literature on their systematic comparison, or their
      application to the square-well fluid
    \item We implement several methods: our own ``simple flat''
      method, Wang-Landau, transition matrix Monte Carlo (TMMC), and a
      hybrid of TMMC with a so-called ``optimized ensemble'' method
    \item Comparison metrics: low energy sampling rates, and relative
      errors in thermodynamic properties computed using results from
      simulating via each method
    \end{itemize}

@@

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.6\columnwidth]{figs/sample-rates.pdf}
      \caption{Iterations per statistically independent energy
        sample.}
      \label{fig:sample_rates}
    \end{figure}
    \begin{figure}[H]
      \centering
      \includegraphics[width=0.6\columnwidth]{figs/u-comps.pdf}
      \caption{Maximum errors in a set of simulations, relative to a
        method which controls for system size and simulation time.}
      \label{fig:errors}
    \end{figure}

@@

  \headerbox{Conclusions}
  {name=conclusions,column=2,below=results}{%

  \headerbox{Acknowledgments}
  {name=thanks,column=2,below=conclusions,above=bottom}{%

>>>>

%%% math and figures
\usepackage{amsmath,graphicx,float}

%% caption text
\usepackage[font=normalsize,labelfont=bf]{caption}

@@

\begin{poster}{
% poster options
    %% colors and appearance
    borderColor=border_color, %
    headerColorOne=header_color_left, %
    headerColorTwo=header_color_right, %
    headerFontColor=header_font_color, %
    boxColorOne=box_color, %
    textborder=rectangle, %
    headerborder=open, %
    background=none, %
    %% headers
    headerfont=\Large\sf\bf, %
    headerheight=0.11\textheight, %
    headershape=roundedright, %
    %% other
    columns=3 %
  }%
  {% top left logo
  }%
  {% title
    \LARGE\title
  }%
  {% authors
    \Large\author \\[1mm]
    \location
  }%
  {% top right logo
    % \includegraphics[width=2.5cm]{figs/osu.pdf}
    % \hspace{5mm}
  }%

@@

    \begin{itemize}
    \item Model system used to capture low order effects of
      short-range attractive forces, e.g. the van Der Waals force
    \item Composed of spheres with diameter $\sigma$ and the pair
      potential given in Figure \ref{fig:pair_potential}
    \item $\lambda$ is called the well width, and $\epsilon$ the well
      depth
    \item First part of potential, $v_{sw}(r<\sigma)=\infty$,
      forbids spheres from overlapping
    \item Second part of potential,
      $v_{sw}(r:\sigma<r<\lambda\sigma)=-\epsilon$, associates an
      energy $-\epsilon$ with each pair of spheres whose centers are
      within distance of $\lambda\sigma$ of each other
    \item Spheres whose centers are more than $\lambda\sigma$ apart do
      not interact
    \end{itemize}

@@

    \begin{itemize}
    \item Model systems are powerful tools for understanding complex
      physical systems, but they do not exist in the real world
    \item Theories of model systems are tested against Monte Carlo
      simulations
    \item Metropolis Monte Carlo fluid simulations: random walk in
      configuration space
    \item Collect histograms of observed system properties
      (e.g. potential energy,
    \item Critical point: highly localized distributions of spheres;
      unlikely to be observed in random samples of all possible fluid
      configurations (second law of thermodynamics)
    \item For the square-well fluid, localized distribution of spheres
      implies a low energy
    \end{itemize}

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.55\columnwidth]
      {figs/dos-poster-example.pdf}
      \caption{Sample density of states of a particular square-well
        fluid ($\lambda=1.3$, $\eta=0.3$), as computed via several
        methods. The logarithmic span of the density of states is
        proportional to the number of moves one must simulate via
        standard Monte Carlo in order to observe the energies with the
        lowest state densities.}
      \label{fig:dos}
    \end{figure}

@@

    \begin{itemize}
    \item Means for biasing Monte Carlo simulations in favor of lower
      energy states
    \item Accept all moves in simulation which lower the system's
      energy, probabilistically reject moves to higher energy
    \item Details of rejection process define a histogram method
    \item Little to no literature on systematic comparison of
      published histogram methods, or on application to the
      square-well fluid
    \item We implement several methods: our own ``simple flat''
      method, Wang-Landau, transition matrix Monte Carlo (TMMC), and a
      hybrid of TMMC with a so-called ``optimized ensemble'' method
    \item Comparison metrics:
      \begin{itemize}
      \item low energy sampling rates
      \item errors in fluid properties computed from simulation
        results
      \end{itemize}
    \end{itemize}

@@

    \begin{itemize}
    \item Errors computed relative to long, ``gold standard''
      simulations
    \end{itemize}

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.6\columnwidth]{figs/sample-rates.pdf}
      \caption{Simulation iterations per independent energy sample.}
      \label{fig:sample_rates}
    \end{figure}
    \begin{figure}[H]
      \centering
      \includegraphics[width=0.5\columnwidth]{figs/u-comps.pdf}
      \caption{Maximum errors in a series of simulations, relative to
        a method which controls for system size and simulation time.}
      \label{fig:errors}
    \end{figure}

    \begin{itemize}
    \item Simple flat method: performs decently in most cases, but has
      a few outlying cases of failure
    \item Wang-Landau: performs well if provided with a predefined
      energy range, otherwise fails catastrophically
    \item TMMC: decent performance; no failures; does not require
      predefined energy range
    \item Optimized ensemble applied to TMMC: worsens performance (by
      all metrics)
    \end{itemize}

@@

  \headerbox{Conclusions}
  {name=conclusions,column=2,below=results,above=bottom}{%

    \begin{itemize}
    \item Simple flat method possibly unreliable, though may be worth
      investigating further (current algorithm is minimalistic)
    \item For systems with a known energy range, or a particular
      energy range of interest, Wang-Landau is best for exploring all
      energies
    \item TMMC: most reliable for systems without a known energy range
    \item Optimized ensemble: designed to optimize standard strategy
      for histogram methods, but only did damage; this strange result
      is worth further investigation
    \end{itemize}
